### 2022-01-11 acks.tex
new approach section

\section*{Acknowledgements}\label{sec:acks}

This research has been partially funded by
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item the VINNOVA Competence Center for \ac{TECoSA} at KTH Royal Institute of Technology
    \item the \ac{SSF}, through grant number ITM17--0246, awarded to James Gross, Ph.D.
\end{enumerate*}
Any opinions, findings, conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the view(s) of their employers or funding sources.

We also wish to thank S. Mostafavi, V. N. Moothedath, M. Alsakati, O. Ferm, S. Vojcic, and A. Bhattacharya for their effort and support for this work.