### 2022-01-12 abstract.tex
neel fixes

\begin{abstract}
    \acfp*{NCS} have long been a subject of theoretical research, however, little work has been devoted to repeatable and scalable experimentation.
    The advent of edge computing and the subsequent proliferation of \acsp*{NCS} has quickly led this to become a key issue that needs to be addressed.

    In this work, we present a novel, completely software-based approach to this challenge.
    Our core idea rests on a framework for the emulation of physical plants which communicate with controllers implemented in software over a real network.
    This approach is furthermore designed and built for the Edge, using a general-purpose programming language and with full compatibility with industry-standard containerization solutions.
    We validate this framework using an initial implementation of an inverted pendulum \acs*{NCS}.
    Our results showcase the utility of the tool as a repeatable, extensible, and scalable solution to \acs*{NCS} performance evaluation.
\end{abstract}