### 2022-01-13 conclusion.tex

\section{Conclusion}\label{sec:conclusion}

The issue of repeatable and scalable benchmarks has been largely glossed over in \ac{NCS} literature, as existing experimental research studies tend to implement \emph{ad-hoc} solutions.
To the best of our knowledge, the work by \textcite{Zoppi2020NCSBench} is the only available framework for \ac{NCS} benchmarking focused on repeatability and reproducibility.

In this work, we aim to tackle this issue by presenting a fully software-based framework for repeatable, reproducible, and easily scalable \ac{NCS} benchmarking with a particular focus on edge deployment.
Our framework, \ac{CLEAVE}, allows for the easy implementation of emulated \acp{NCS} on real networks.
It is fully parametrizable, extendable, and built with industry-standard edge technologies and paradigms in mind.
We validate the utility of this tool through a series of scenarios; our results showcase the ability of the framework to extract relevant metrics relating to the stability of the control system, as well as on the perfomance of the underlying network link.
We believe \ac{CLEAVE} represents an important step towards enabling inexpensive and low-complexity scalable research for real-world deployment of edge-bound \acp{NCS}.

There is still, however, work to be done.
The current implementation only includes a single plant-controller pair; we envision creating an open library of \acp{NCS} to share with the community.
At the moment, the interactions of \ac{CLEAVE} and tools such as Docker are still quite superficial.
Our goal is to achieve a much tighter integration, e.g.\ by providing the toolkit as pre-packaged container images.
Finally, the validity of the results obtained by the framework will have to be verified through more thorough, realistic scenarios than what we have been able to show in this work.
In particular, we intend to perform large-scale experimentation targetting 5G cellular deployments, as this technology is set to become the backbone of edge networks in the near future.