### 2022-02-11 abstract.tex
fix abstract

\begin{abstract}
    As the number of cyber-physical systems rises, it becomes increasingly crucial to study \acfp*{NCS} combining control communication co-design. This nature of \acsp*{NCS} has led to task-specific approaches to research, creating a dearth of generalizable, repeatable, and scalable experimentation. Further, with the advent of edge computing solutions, it is of paramount importance to explore its relevance in such applications. In this work, we present \acs*{CLEAVE}, a novel, completely software-based framework for repeatable and scalable experimentation in edge native \acsp*{NCS}.
    Our approach is based on the emulation of physical plants communicating over a real network with software-based controllers.
    \acs*{CLEAVE} is designed and built for the edge, using Python3 and with full compatibility with industry-standard containerization solutions.
    Although designed for single-loop emulations, the flexibility afforded by the aforementioned characteristics allow our framework to be adapted to a multitude of complex scenarios.

    We validate \acs*{CLEAVE} using an initial implementation of an inverted pendulum \acs*{NCS}.
    Our results showcase the utility of the tool as a repeatable, extensible, and scalable solution to \acs*{NCS} performance evaluation and benchmarking on the Edge.
\end{abstract}